---
title: "Central Limit Theorem"
author: Corvus LEE
date: 2019-12-28
output:
  github_document:
    toc: yes
---

```{r include=FALSE}
# backup system default par
opar <- par(no.readonly = TRUE)

# define my default par
default_par <- function(){
  par(opar)
  par(col='grey', col.axis='grey', col.lab='grey', col.main='grey', col.sub='grey')
}
```
# 1. What is CLT?

If the sample size is sufficiently large enough, then the mean of the samples will follow a normal distribution, no matter what the underlying population distribution is. In this notebook I will illustrate how well the theorem applies to different distributions with different sample sizes.

# 2. Population: Normal distribution

```{r}
# Generate a standard normal population with size 10000
set.seed(8000)
y <- rnorm(10000)
```

```{r echo=FALSE}
# Visualize the population distribution
default_par()
par(mfrow=c(1,2))

# Histogram
hist(y, col='dodgerblue', ann=FALSE, xaxt='n', yaxt='n', border='white')
title(xlab='population distribution')
axis(1, col='grey')
abline(v=mean(y), col='black', lty=2)
text(x=2.2, y=1700, paste('mean=',round(mean(y),2)), col='black')

# QQ plot
qqnorm(y, col='dodgerblue', ann=FALSE, xaxt='n', yaxt='n', frame.plot=FALSE)
title(xlab='normal quantiles', ylab='population quantiles')
axis(1, col='grey')
axis(2, col='grey')
qqline(y, col='black')
```

Sample mean will always follow normal distribution for all sample sizes

```{r}
# Generate 500 samples with sample size of one
sample <- double()
set.seed(8001)
for (i in 1:500) {
  sample <- append(sample, sample(y,1))
}
```

```{r echo=FALSE}
# Visualize the sample distribution
default_par()
par(mfrow=c(1,2))

# Histogram
hist(sample, col='orange', ann=FALSE, xaxt='n', yaxt='n', border='white')
title(xlab='sample distribution')
axis(1, col='grey')
abline(v=mean(sample), col='black', lty=2)
text(x=1.4, y=90, paste('mean=',round(mean(sample),2)), col='black')

# QQ plot
qqnorm(sample, col='orange', ann=FALSE, xaxt='n', yaxt='n', frame.plot=FALSE)
title(xlab='normal quantiles', ylab='sample quantiles')
axis(1, col='grey')
axis(2, col='grey')
qqline(sample, col='black')

# Title
mtext("Sample mean distribution",
      col='#666666', line=-1, outer = TRUE, cex = 1.2, adj = 0)
mtext("Sample size: 1",
      col='grey', line=-2, outer=TRUE, adj=0)
```

# 3. Population: Uniform distribution

```{r}
# Generate an uniform population with size 10000
set.seed(8000)
y <- runif(10000)
```

```{r echo=FALSE}
# Visualize the population distribution
default_par()

# Histogram
hist(y, col='dodgerblue', ann=FALSE, xaxt='n', yaxt='n', border='white')
title(xlab='population distribution')
axis(1, col='grey')
abline(v=mean(y), col='black', lty=2)
text(x=.6, y=550, paste('mean=',round(mean(y),2)), col='black')
```

CLT applies well to uniformly distributed population with small sample size (e.g., 5)

```{r}
# Generate 500 samples with various sample sizes
sample1 <- double()
sample5 <- double()

# sample size: 1
set.seed(8001)
for (i in 1:500) {
  sample1 <- append(sample1, mean(sample(y,1)))
}

# sample size: 5
set.seed(8001)
for (i in 1:500) {
  sample5 <- append(sample5, mean(sample(y,5)))
}
```

```{r echo=FALSE, fig.asp=10/9}
# Visualize the sample distribution
default_par()
par(mfrow=c(2,2))

# sample size 1: histogram
hist(sample1, col='grey', ann=FALSE, xaxt='n', yaxt='n', border='white')
title(xlab='sample distribution')
axis(1, col='grey')
abline(v=mean(sample1), col='black', lty=2)
text(x=.7, y=50, paste('mean=',round(mean(sample1),2)), col='black')

# sample size 1: QQ plot
qqnorm(sample1, col='grey', ann=FALSE, xaxt='n', yaxt='n', frame.plot=FALSE)
title(xlab='normal quantiles', ylab='sample quantiles')
axis(1, col='grey')
axis(2, col='grey')
qqline(sample1, col='black')

# sample size 5: histogram
hist(sample5, col='orange', ann=FALSE, xaxt='n', yaxt='n', border='white')
title(xlab='sample distribution')
axis(1, col='grey')
abline(v=mean(sample5), col='black', lty=2)
text(x=.7, y=50, paste('mean=',round(mean(sample5),2)), col='black')

# sample size 5: QQ plot
qqnorm(sample5, col='orange', ann=FALSE, xaxt='n', yaxt='n', frame.plot=FALSE)
title(xlab='normal quantiles', ylab='sample quantiles')
axis(1, col='grey')
axis(2, col='grey')
qqline(sample5, col='black')

# Title
mtext("Sample mean distributions",
      col='#666666', line=-1.2, outer = TRUE, cex = 1.2, adj = 0)
mtext("Sample size: 1, 5",
      col='grey', line=-2.2, outer=TRUE, adj=0)
```

# 4. Population: Exponential distribution

```{r}
# Generate an exponential population with size 10000
set.seed(8000)
y <- rexp(10000)
```

```{r echo=FALSE}
# Visualize the population distribution
default_par()

# Histogram
hist(y, col='dodgerblue', ann=FALSE, xaxt='n', yaxt='n', border='white')
title(xlab='population distribution')
axis(1, col='grey')
abline(v=mean(y), col='black', lty=2)
text(x=2, y=1700, paste('mean=',round(mean(y),2)), col='black')
```

With an exponential distribution, CLT starts to apply with a sample size of ten. The distribution looks much nicer when we have a sample size of a hundred.

```{r}
# Generate 500 samples with various sample sizes
sample5 <- double()
sample10 <- double()
sample100 <- double()

# sample size: 5
set.seed(8001)
for (i in 1:500) {
  sample5 <- append(sample5, mean(sample(y,5)))
}

# sample size: 10
set.seed(8001)
for (i in 1:500) {
  sample10 <- append(sample10, mean(sample(y,10)))
}

# sample size: 100
set.seed(8001)
for (i in 1:500) {
  sample100 <- append(sample100, mean(sample(y,100)))
}
```

```{r echo=FALSE, fig.asp=3/2}
# Visualize the sample distributions
default_par()
par(mfrow=c(3,2))

# sample size 5: histogram
hist(sample5, col='grey', ann=FALSE, xaxt='n', yaxt='n', border='white')
title(xlab='sample distribution')
axis(1, col='grey')
abline(v=mean(sample5), col='black', lty=2)
text(x=1.5, y=150, paste('mean=',round(mean(sample5),2)), col='black')

# sample size 5: QQ plot
qqnorm(sample5, col='grey', ann=FALSE, xaxt='n', yaxt='n', frame.plot=FALSE)
title(xlab='normal quantiles', ylab='sample quantiles')
axis(1, col='grey')
axis(2, col='grey')
qqline(sample5, col='black')

# sample size 10: histogram
hist(sample10, col='grey', ann=FALSE, xaxt='n', yaxt='n', border='white')
title(xlab='sample distribution')
axis(1, col='grey')
abline(v=mean(sample10), col='black', lty=2)
text(x=1.3, y=110, paste('mean=',round(mean(sample10),2)), col='black')

# sample size 10: QQ plot
qqnorm(sample10, col='grey', ann=FALSE, xaxt='n', yaxt='n', frame.plot=FALSE)
title(xlab='normal quantiles', ylab='sample quantiles')
axis(1, col='grey')
axis(2, col='grey')
qqline(sample10, col='black')

# sample size 100: histogram
hist(sample100, col='orange', ann=FALSE, xaxt='n', yaxt='n', border='white')
title(xlab='sample distribution')
axis(1, col='grey')
abline(v=mean(sample100), col='black', lty=2)
text(x=1.1, y=90, paste('mean=',round(mean(sample100),2)), col='black')

# sample size 100: QQ plot
qqnorm(sample100, col='orange', ann=FALSE, xaxt='n', yaxt='n', frame.plot=FALSE)
title(xlab='normal quantiles', ylab='sample quantiles')
axis(1, col='grey')
axis(2, col='grey')
qqline(sample100, col='black')

# Title
mtext("Sample mean distributions",
      col='#666666', line=-1.3, outer = TRUE, cex = 1.2, adj = 0)
mtext("Sample size: 5, 10, 100",
      col='grey', line=-2.3, outer=TRUE, adj=0)
```

# 5. Population: Other distribution

```{r}
# Generate a population with size 10000
set.seed(8000)
y <- rbeta(10000, .5, .5)
```

```{r echo=FALSE}
# Visualize the population distribution
default_par()

# Histogram
hist(y, col='dodgerblue', ann=FALSE, xaxt='n', yaxt='n', border='white')
title(xlab='population distribution')
axis(1, col='grey')
abline(v=mean(y), col='black', lty=2)
text(x=.6, y=500, paste('mean=',round(mean(y),2)), col='black')
```

Surprisingly CLT applies really well to this beta distribution

```{r}
# Generate 500 samples with various sample sizes
sample5 <- double()

# sample size: 5
set.seed(8001)
for (i in 1:500) {
  sample5 <- append(sample5, mean(sample(y,5)))
}

```

```{r echo=FALSE}
# Visualize the sample distributions
default_par()
par(mfrow=c(1,2))

# sample size 5: histogram
hist(sample5, col='orange', ann=FALSE, xaxt='n', yaxt='n', border='white')
title(xlab='sample distribution')
axis(1, col='grey')
abline(v=mean(sample5), col='black', lty=2)
text(x=.7, y=50, paste('mean=',round(mean(sample5),2)), col='black')

# sample size 5: QQ plot
qqnorm(sample5, col='orange', ann=FALSE, xaxt='n', yaxt='n', frame.plot=FALSE)
title(xlab='normal quantiles', ylab='sample quantiles')
axis(1, col='grey')
axis(2, col='grey')
qqline(sample5, col='black')

# Title
mtext("Sample mean distribution",
      col='#666666', line=-1, outer = TRUE, cex = 1.2, adj = 0)
mtext("Sample size: 5",
      col='grey', line=-2, outer=TRUE, adj=0)
```
